This appendix describes the calibration strategy for the computational implementation of the model. The current implementation covers the core features of the theoretical framework (see Section~\ref{s:model}); features not yet implemented are listed in Section~\ref{sec:not-implemented}.

\subsection{Model Overview}

The implemented model is an OLG model with heterogeneous agents featuring:
\begin{enumerate}
    \item Lifecycle structure: $J$ periods, exogenous retirement at $J_R$, stochastic survival $\pi(j)$.
    \item Idiosyncratic risk: stochastic income ($n_y$ states, Tauchen AR(1)) and employment. Health expenditure enters as a fixed per-capita cost, not as an idiosyncratic state.
    \item Labor supply: intensive margin ($u(c,\ell)$) or extensive margin (binary employment), configurable.
    \item Education heterogeneity: low, medium, high types with education-specific income processes and unemployment rates.
    \item Perfect foresight over aggregates: agents know the full path of $\{r_t, w_t, K_t^g\}$.
    \item Government sector: taxes $(\tau^c, \tau^l, \tau^p, \tau^k, \tau^{\text{beq}})$, UI, pensions, health subsidies, government consumption $G_t$, and public investment $I_t^g$.
    \item Production: Cobb--Douglas with public capital $Y_t = A(K_t^g)^{\eta_g} K_t^\alpha L_t^{1-\alpha}$.
    \item Demographics: time-varying population growth $g_N$.
\end{enumerate}

\subsection{Current Parameter Values}\label{sec:current-params}

Table~\ref{tab:current-params} lists the parameter values used in the fiscal experiments of Section~\ref{s:experiments}.

\begin{table}[H]
\centering
\small
\caption{Parameter values used in the fiscal experiment.}
\label{tab:current-params}
\begin{tabular}{llll}
\toprule
Parameter & Description & Value & Source \\
\midrule
\multicolumn{4}{l}{\textit{Lifecycle and preferences}} \\
$J$ & Lifecycle length & 60 & \\
$J_R$ & Retirement period & 45 & \\
$\beta$ & Discount factor & 0.96 & \\
$\sigma$ & Risk aversion (CRRA) & 2.0 & \\
$\underline{a}$ & Borrowing constraint & 0 & \\
\midrule
\multicolumn{4}{l}{\textit{Production}} \\
$\alpha$ & Private capital share & 0.33 & National accounts \\
$\delta$ & Private capital depreciation & 0.05 & \\
$A$ & TFP & 1.0 & Normalization \\
$\eta_g$ & Public capital output elasticity & 0.10 & \cite{BaxterKing_AER93} \\
$\delta_g$ & Public capital depreciation & 0.05 & \\
$K_0^g$ & Initial public capital & 1.0 & \\
\midrule
\multicolumn{4}{l}{\textit{Tax rates (fiscal experiment baseline)}} \\
$\tau^l$ & Labour income tax & 0.15 & \\
$\tau^c$ & Consumption tax & 0.18 & \\
$\tau^p$ & Payroll tax & 0.20 & \\
$\tau^k$ & Capital income tax & 0.20 & \\
$\rho$ & Pension replacement rate & 0.60 & \\
$\rho^{ui}$ & UI replacement rate & 0.30 & \\
\midrule
\multicolumn{4}{l}{\textit{Health expenditure}} \\
$m$ & Per-capita medical cost (share of mean income) & 0.05 & OECD Health Statistics \\
$\kappa$ & Government health coverage rate & 0.90 & OECD Health Statistics \\
\midrule
\multicolumn{4}{l}{\textit{Demographics and experiment}} \\
$g_N$ & Population growth rate & 0.02 & \\
$r$ & Interest rate (exogenous) & 0.04 & \\
$T_{trans}$ & Transition horizon & 40 & \\
$B_0$ & Initial government debt & 0.0 & \\
\bottomrule
\end{tabular}
\end{table}

\subsection{Externally Calibrated Parameters}

These parameters are set directly from data or existing literature estimates.

\paragraph{Preferences and lifecycle.}

\begin{table}[H]
\centering
\small
\begin{tabular}{llll}
\toprule
Parameter & Description & Source & Range \\
\midrule
$J$ & Lifecycle length & Life expectancy $-$ entry age & 40--60 \\
$J_R$ & Retirement age & Statutory/effective retirement & 30--45 \\
$\sigma$ & Risk aversion (CRRA) & Chetty (2006), Attanasio \& Weber & 1--4 \\
$\underline{a}$ & Borrowing constraint & Institutional & $\geq 0$ \\
\bottomrule
\end{tabular}
\end{table}

\paragraph{Production.}

\begin{table}[H]
\centering
\small
\begin{tabular}{llll}
\toprule
Parameter & Description & Source & Range \\
\midrule
$\alpha$ & Capital share & National accounts & 0.33 \\
$\delta$ & Depreciation rate & Investment/capital ratio & 0.05--0.10 \\
\bottomrule
\end{tabular}
\end{table}

\paragraph{Tax rates.}
Tax rates can be time-varying paths $\{\tau^c_t, \tau^l_t, \tau^p_t, \tau^k_t\}$ for transition experiments.

\begin{table}[H]
\centering
\small
\begin{tabular}{llll}
\toprule
Parameter & Description & Source & Range \\
\midrule
$\tau^c$ & Consumption tax & Effective rate from fiscal data & 0.05--0.25 \\
$\tau^l$ & Labor income tax & Effective rate from fiscal data & 0.10--0.30 \\
$\tau^p$ & Payroll tax (wages only) & Social security contributions & 0.10--0.25 \\
$\tau^k$ & Capital income tax & Effective rate from fiscal data & 0.15--0.30 \\
\bottomrule
\end{tabular}
\end{table}

\paragraph{Transfers.}

\begin{table}[H]
\centering
\small
\begin{tabular}{llll}
\toprule
Parameter & Description & Source & Range \\
\midrule
$\rho^{\text{pens}}$ & Pension replacement rate & SS replacement rate data & 0.40--0.80 \\
$\rho^{\text{ui}}$ & UI wage replacement rate & Statutory UI rules & 0.20--0.60 \\
\bottomrule
\end{tabular}
\end{table}

\paragraph{Health expenditure.}

\begin{table}[H]
\centering
\small
\begin{tabular}{llll}
\toprule
Parameter & Description & Source & Range \\
\midrule
$m$ & Per-capita medical cost (share of mean income) & OECD Health Statistics / Eurostat & 0.05--0.15 \\
$\kappa$ & Govt.\ coverage rate of medical costs & OECD Health Statistics / Eurostat & 0.50--0.90 \\
\bottomrule
\end{tabular}
\end{table}

\noindent $m$ is set to match total health expenditure as a share of GDP in the calibrated country. $\kappa$ is the public share of total health expenditure. Both are available from OECD Health Statistics and Eurostat (SHA framework).

\paragraph{Education and demographics.}

\begin{table}[H]
\centering
\small
\begin{tabular}{llll}
\toprule
Parameter & Description & Source & Range \\
\midrule
Education shares & Population shares by education & Census / Eurostat & Country-specific \\
$\mu_z^{(e)}$ & Mean log income by education $e$ & EU-SILC via LIS, see \S\ref{sec:lis-income} & Education-specific \\
$\rho_z^{(e)}$ & Income persistence by education $e$ & EU-SILC via LIS, see \S\ref{sec:lis-income} & Education-specific \\
$\sigma_\eta^{(e)}$ & Innovation std.\ dev.\ by education $e$ & EU-SILC via LIS, see \S\ref{sec:lis-income} & Education-specific \\
$u^{(e)}$ & Unemployment rate by education $e$ & Eurostat & Education-specific \\
$g_N$ & Population growth rate & Demographic projections & $-0.01$--$0.02$ \\
\bottomrule
\end{tabular}
\end{table}

\subsubsection{Income Process from LIS Greek EU-SILC}\label{sec:lis-income}

The AR(1) labor productivity process per education stratum is calibrated externally using Luxembourg Income Study (LIS) microdata for Greece, accessed via LISSY remote execution. Greek EU-SILC in LIS is cross-sectional only (no cross-wave personal identifier), so persistence cannot be identified from within-individual autocovariances; identification of $\sigma_\eta$ relies on the cross-sectional age profile of $\mathrm{Var}(\log y)$, with $\rho$ fixed at a baseline value and treated as a sensitivity parameter.

\paragraph{Data.} 19 Greek EU-SILC waves pooled (GR03--GR21). Master sample after age 25--60 and valid-education filters: 222{,}246 observations. Primary specification (\textsf{emp\_fyft}) selects employees by $\textsf{pi11} > 0$ (positive wage income) and $\textsf{fyft} = 1$ (full-year-full-time). Top/bottom 0.5\% of $\log y$ trimmed within each (education, year) cell; residualization on year fixed effects within education stratum.\footnote{Greek EU-SILC populates LIS \textsf{status1} only with codes $\{100, 210, 220, 240\}$, with no $\{110, 120\}$ split, so a \textsf{status1}-based employee filter returns zero observations. The $\textsf{pi11} > 0$ criterion is the cross-country-robust alternative.}

\begin{table}[H]
\centering
\small
\caption{LIS Greek pool: sample composition by education (\textsf{emp\_fyft}, GR03--GR21).}
\label{tab:lis-sample}
\begin{tabular}{lccc}
\toprule
Education & $N$ obs & Mean $\log y$ & SD $\log y$ \\
\midrule
Low (ISCED 0--2)    & 11{,}718 & 9.489 & 0.432 \\
Medium (ISCED 3--4) & 32{,}211 & 9.642 & 0.396 \\
High (ISCED 5+)     & 32{,}326 & 9.901 & 0.443 \\
\midrule
Total              & 76{,}255 &       &       \\
\bottomrule
\end{tabular}
\end{table}

\paragraph{Variance profile.} Table~\ref{tab:lis-variance-profile} reports the cross-sectional variance of residualized log earnings, $\mathrm{Var}(u_j)$, by 5-year age band and education stratum. The age slope of $\mathrm{Var}(u_j)$ is the moment that identifies $\sigma_\eta$ (given $\rho$).

\begin{table}[H]
\centering
\small
\caption{$\mathrm{Var}(u_j)$ by 5-year age band and education (\textsf{emp\_fyft}).}
\label{tab:lis-variance-profile}
\begin{tabular}{lccc}
\toprule
Age band & Low & Medium & High \\
\midrule
25--29 & 0.160 & 0.091 & 0.115 \\
30--34 & 0.136 & 0.102 & 0.129 \\
35--39 & 0.140 & 0.115 & 0.143 \\
40--44 & 0.178 & 0.124 & 0.155 \\
45--49 & 0.166 & 0.136 & 0.152 \\
50--54 & 0.157 & 0.143 & 0.150 \\
55--59 & 0.168 & 0.148 & 0.158 \\
\bottomrule
\end{tabular}
\end{table}

\paragraph{Estimation strategy.} Within each education stratum we fit
\begin{equation}\label{eq:lis-md}
\mathrm{Var}(u_j) = \sigma_\alpha^2 + \sigma_\eta^2 \cdot \frac{1 - \rho^{2j}}{1 - \rho^2},
\end{equation}
where $j = \text{age} - 25$ is years since the model anchor age. Estimation is by weighted nonlinear least squares (weights = cell sample sizes) on the 36 single-year-age cells per education. We report two specifications: (i) a joint fit of $(\rho, \sigma_\eta, \sigma_\alpha)$, and (ii) the calibration baseline with $\rho = 0.95$ and $(\sigma_\eta, \sigma_\alpha)$ estimated. Specification (ii) is the baseline: $\rho$ is weakly identified from cross-sectional moments alone, and the joint specification is reported as sensitivity. The permanent-component estimate $\sigma_\alpha$ is informational---the calibrated model has no fixed effect, so $\sigma_\alpha$ loads onto education-type mean dispersion via $\mu_z^{(e)}$.

\begin{table}[H]
\centering
\small
\caption{Estimated AR(1) parameters by education, LIS Greek pool.}
\label{tab:lis-estimates}
\begin{tabular}{lccc}
\toprule
 & Low & Medium & High \\
\midrule
\multicolumn{4}{l}{\textit{Joint MD: $(\rho, \sigma_\eta, \sigma_\alpha)$}} \\
\quad $\rho$ & 0.961 & 0.980 & 0.954 \\
\quad $\sigma_\eta$ & 0.047 & 0.061 & 0.072 \\
\quad $\sigma_\alpha$ & 0.370 & 0.280 & 0.320 \\
\midrule
\multicolumn{4}{l}{\textit{Fixed $\rho = 0.95$ (baseline): $(\sigma_\eta, \sigma_\alpha)$}} \\
\quad $\sigma_\eta$ & 0.054 & 0.086 & 0.075 \\
\quad $\sigma_\alpha$ & 0.367 & 0.259 & 0.318 \\
\midrule
$\mu_z^{(e)} - \mu_z^{(\text{med})}$ & $-0.153$ & $0.000$ & $+0.259$ \\
\bottomrule
\end{tabular}
\end{table}

\begin{table}[H]
\centering
\small
\caption{Sensitivity of $\sigma_\eta^{(e)}$ to fixed $\rho$.}
\label{tab:lis-rho-sensitivity}
\begin{tabular}{lccc}
\toprule
$\rho$ & Low & Medium & High \\
\midrule
0.90 & 0.073 & 0.118 & 0.110 \\
0.95 & 0.054 & 0.086 & 0.075 \\
0.97 & 0.042 & 0.070 & 0.059 \\
0.99 & 0.030 & 0.053 & 0.043 \\
\bottomrule
\end{tabular}
\end{table}

\paragraph{Comparison with proxies.} Provisional values from Cooper, Haan \& Zhu (BIS WP 1102) for Italy/Spain are $\rho \approx 0.95$ and $\sigma_\eta \approx 0.13$. The Greek estimates at $\rho = 0.95$ are $\sigma_\eta \in \{0.054, 0.086, 0.075\}$---half or less of the Italy/Spain proxy. The flatter Greek age profile of $\mathrm{Var}(u_j)$ is consistent with stronger labour-market institutions (collective bargaining coverage, large public-sector share) and with selection into formal full-year employment among the \textsf{emp\_fyft} sample.

\subsection{Internally Calibrated Parameters}\label{sec:internal-cal}

The income process $(\rho_z^{(e)}, \sigma_\eta^{(e)}, \mu_z^{(e)})$ is pinned externally from LIS estimation (\S\ref{sec:lis-income}); job separation rates and per-education unemployment rates are likewise pinned from external data. Two parameters are calibrated internally via the Simulated Method of Moments (SMM).

\begin{table}[H]
\centering
\small
\begin{tabular}{lll}
\toprule
Parameter & Role & Identified by \\
\midrule
$\beta$ & Discount factor & Wealth-to-income ratio $A/Y$ \\
$\nu$ & Frisch labour-disutility scale & Average hours worked \\
\bottomrule
\end{tabular}
\end{table}

\paragraph{Notes.}
\begin{itemize}[leftmargin=*]
    \item In the small-open-economy specification the interest rate $r$ is exogenous, so $\beta$ is identified by the household-saving target $A/Y$ rather than the closed-economy capital-output ratio $K/Y$.
    \item TFP $A$ is set to a normalization ($A = 1$) and the wage is recovered from the firm's first-order condition; no calibration of $A$ or $Z$ is needed.
    \item The job separation rate is not a free parameter---it is derived from $u^{(e)}$ and $\lambda^{\text{find}}$: $\lambda^{\text{sep}} = \min\!\big(\frac{u}{1-u}\lambda^{\text{find}},\;\bar{\lambda}^{\text{sep}}\big)$.
    \item Income-process parameters are excluded from the SMM problem because $\rho$ is weakly identified from cross-sectional moments alone (\S\ref{sec:lis-income}). Folding $\rho$ and $\sigma_\eta$ into the SMM loop would let wealth-distribution residuals contaminate parameters that have a clean external identification from LIS.
\end{itemize}

\subsection{Target and Validation Moments}\label{sec:targets}

The SMM problem targets two macro moments. The remaining moments listed below are reported as untargeted validation moments---they discipline the calibration ex post but are not in the objective function.

\begin{table}[H]
\centering
\small
\begin{tabular}{lll}
\toprule
Moment & Data source & Identifies \\
\midrule
\multicolumn{3}{l}{\textit{Targeted (in SMM objective)}} \\
Wealth-to-income ratio $A/Y$ & National accounts & $\beta$ \\
Average hours worked & Labour-force survey & $\nu$ \\
\midrule
\multicolumn{3}{l}{\textit{Validation: wealth and income distribution}} \\
Wealth Gini coefficient & HFCS & income process, $\underline{a}$ \\
Income Gini coefficient & EU-SILC & income process \\
Fraction with zero/negative wealth & HFCS & $\underline{a}$, income process \\
Median wealth-to-income ratio & HFCS & $\beta$ \\
$p_{90}/p_{10}$ income & EU-SILC & income process \\
\midrule
\multicolumn{3}{l}{\textit{Validation: income process (LIS estimation, \S\ref{sec:lis-income})}} \\
Mean $\mathrm{Var}(\log y)$ across ages & EU-SILC via LIS & $\sigma_\eta$, $\sigma_\alpha$ \\
Slope of $\mathrm{Var}(\log y)$ in age & EU-SILC via LIS & $\sigma_\eta$, $\rho$ \\
\midrule
\multicolumn{3}{l}{\textit{Validation: labour market}} \\
Aggregate unemployment rate & Eurostat & $u^{(e)}$, $\lambda^{\text{find}}$ \\
\midrule
\multicolumn{3}{l}{\textit{Validation: fiscal}} \\
Tax revenue / GDP & Eurostat & Tax rates (external) \\
Government spending / GDP & Eurostat & Pension/health (external) \\
\bottomrule
\end{tabular}
\end{table}

\paragraph{Notes.}
\begin{itemize}[leftmargin=*]
    \item Wealth Gini and income Gini are kept as validation rather than targets because the income process is pinned externally from LIS, so a wealth-Gini residual cannot ``correct'' $\sigma_\eta$ or $\rho$ without contaminating the clean external identification (\S\ref{sec:internal-cal}).
    \item The borrowing constraint $\underline{a}$ is set institutionally to zero; conditional on that, the wealth-distribution moments are predictions of the calibrated model rather than objects of fit.
\end{itemize}

\subsection{Calibration Algorithm}

\subsubsection{SMM Objective Function}

With the income process and tax/transfer policy parameters pinned externally (\S\ref{sec:lis-income}), and with the interest rate $r$ exogenous in the small-open-economy specification, the calibration reduces to a two-parameter SMM problem in $\theta = (\beta, \nu)$:
\begin{equation}\label{eq:smm-objective}
Q(\theta) = \big[m^{\text{data}} - m^{\text{model}}(\theta)\big]'\, W\, \big[m^{\text{data}} - m^{\text{model}}(\theta)\big],
\end{equation}
with target vector $m^{\text{data}} = (A/Y,\, \overline{h})$ and identity weighting matrix $W$. There is no GE iteration: the wage is recovered from the firm's first-order condition once $r$ is given, and household saving (and hence $A/Y$) is matched by varying $\beta$ directly.

\subsubsection{Optimizer}

We use derivative-free methods, as the objective is noisy (simulation-based) and non-smooth (grid search in the inner loop):
\begin{itemize}[leftmargin=*]
    \item \textbf{Nelder--Mead simplex} on logit-transformed bounded parameters is the default for the two-parameter problem.
    \item \textbf{Differential evolution} followed by Nelder--Mead polishing is available as a global-search option for sensitivity exercises.
\end{itemize}

\subsubsection{Simulation Requirements}

\begin{itemize}[leftmargin=*]
    \item $n_{\text{sim}} = 10{,}000$ agents per education type.
    \item Random seeds are fixed across objective evaluations to keep the optimization landscape smooth.
    \item Lifecycle solution and simulation use the JAX backend; aggregation runs on the NumPy/Numba path.
\end{itemize}

\subsubsection{Convergence Criteria}

\begin{itemize}[leftmargin=*]
    \item $Q(\theta) < \text{tol}$ or parameter changes $< 10^{-5}$ across simplex iterations.
    \item Each targeted moment within 5\% of its data value at the optimum, or within 2 standard errors of the data estimate.
\end{itemize}

\subsubsection{Practical Workflow}

\begin{enumerate}
    \item Fix externally calibrated parameters from country-level data: tax/transfer rates, demographics, education shares, health-expenditure shares.
    \item Estimate the AR(1) income process per education stratum from LIS Greek EU-SILC by minimum distance on the cross-sectional variance profile (\S\ref{sec:lis-income}); write $(\rho_z^{(e)}, \sigma_\eta^{(e)}, \mu_z^{(e)})$ into the model configuration.
    \item Run SMM on $(\beta, \nu)$ against $(A/Y,\, \overline{h})$ using the JAX backend.
    \item Report untargeted validation moments: wealth Gini, income Gini, $p_{90}/p_{10}$, mean and slope of $\mathrm{Var}(\log y)$ by age, fiscal ratios.
    \item Sensitivity analysis: vary the LIS-baseline $\rho$ over $\{0.90, 0.95, 0.97, 0.99\}$ (Table~\ref{tab:lis-rho-sensitivity}) and re-run the SMM step; vary risk aversion $\sigma$ and the borrowing constraint $\underline{a}$ to check robustness of the validation moments.
\end{enumerate}

\subsection{Missing Model Features (Implementation To-Do)}

The following features are present in the theoretical model (Section~\ref{s:model}) but not yet in the code. See Section~\ref{sec:not-implemented} for details. They are listed roughly in order of priority.

\begin{enumerate}
    \item \textbf{Continuous human capital accumulation.} The theory specifies $\log h_{j+1} = \log h_j + g_j + \epsilon_j$; the code uses a deterministic age profile $\kappa_j$ with no human capital state.
    \item \textbf{Schooling phase and children.} Child consumption costs $c_y(j)$ and private education expenditure $e(j) = (1-\kappa^{\text{school}}_t)\bar{e}(j)$ are not in the code.
    \item \textbf{Endogenous interest rate (SOE equilibrium).} The theory determines $r_t$ via risk-neutral external lenders at $r^*$; the code takes $\{r_t\}$ as an exogenous input.
    \item \textbf{CES capital aggregator.} The CES effective capital stock $\tilde{K} = [(1-\theta)K^\rho + \theta(K^g)^\rho]^{1/\rho}$ is not implemented; the code uses only Cobb--Douglas.
    \item \textbf{Full earnings-history pension.} The theory uses $\text{PENS} = \max\{\rho\,\bar{y},\,b_{\min}\}$ with $\bar{y}$ as a continuous state; the code approximates with a weighted combination of $z_{\text{last}}$ and mean income.
    \item \textbf{Age- and health-dependent income transitions.} The theory specifies $P_z(z'|z,j,s)$; the code uses a constant $P_z(z'|z)$.
\end{enumerate}

\subsubsection{Additional Parameters Required}

Table~\ref{tab:future-params} lists the parameters that would be added once the missing features are implemented.

\begin{table}[H]
\centering
\small
\caption{Parameters required by missing model features.}
\label{tab:future-params}
\begin{tabular}{lllll}
\toprule
Feature & Parameter & Description & Calibration & Range \\
\midrule
\multicolumn{5}{l}{\textit{1. Human capital accumulation}} \\
& $A_0$ & Initial productivity level & Internal (SMM) & --- \\
& $\gamma^h$ & Returns to education investment & External & 0.3--0.7 \\
& $g_j$ & Deterministic HC growth by age & External & Age profile \\
& $\sigma_\epsilon$ & HC shock std.\ dev. & Internal (SMM) & 0.05--0.20 \\
\midrule
\multicolumn{5}{l}{\textit{2. Schooling and children}} \\
& $c_y(j)$ & Child consumption cost by age & External & 0.1--0.4 \\
& $e(j)$ & Education expenditure by age & External & Data \\
& $\kappa^{\text{school}}_t$ & Govt.\ education subsidy rate & External & 0.5--1.0 \\
& $\mathcal{S}$ & Number of schooling periods & External & 3--6 \\
\midrule
\multicolumn{5}{l}{\textit{3. Sovereign debt and SOE}} \\
& $r^*$ & World interest rate & External & 0.01--0.04 \\
& $B_0$ & Initial govt.\ debt stock & External & Data \\
\midrule
\multicolumn{5}{l}{\textit{4. CES capital aggregator}} \\
& $\theta$ & CES share parameter & External & $\eta_g/(\alpha+\eta_g)$ \\
& $\sigma^{\text{CES}}$ & Elasticity of substitution $K$/$K^g$ & External & 0.5--$\infty$ \\
\midrule
\multicolumn{5}{l}{\textit{5. Earnings-history pension}} \\
& $b_{\min}$ & Minimum pension floor & External & Data \\
& Indexation rule & Price vs.\ wage indexation & External & Policy choice \\
\midrule
\multicolumn{5}{l}{\textit{6. State-dependent income transitions}} \\
& $P_z(z'|z,j,s)$ & Transition probs.\ by age, health & Internal (SMM) & Panel data \\
\bottomrule
\end{tabular}
\end{table}

\subsubsection{Additional Data Requirements}

Table~\ref{tab:future-data} lists the data sources needed to calibrate or validate the parameters above.

\begin{table}[H]
\centering
\small
\caption{Data required for missing model features.}
\label{tab:future-data}
\begin{tabular}{lll}
\toprule
Feature & Data needed & Source \\
\midrule
\multicolumn{3}{l}{\textit{1. Human capital accumulation}} \\
& Wage--age profiles by education & EU-SILC / PSID \\
& Within-cohort earnings dispersion by age & EU-SILC / PSID \\
& Returns to education (Mincer coefficients) & Literature \\
\midrule
\multicolumn{3}{l}{\textit{2. Schooling and children}} \\
& Fertility rates and household size by age & Eurostat / Census \\
& Public and private education spending & COFOG / OECD EAG \\
& Childcare costs by age of child & OECD Family Database \\
\midrule
\multicolumn{3}{l}{\textit{3. Sovereign debt and SOE}} \\
& Government debt stock and maturity profile & Eurostat / ECB \\
& Sovereign bond yields & ECB / Bloomberg \\
& Net international investment position & Eurostat / IMF \\
\midrule
\multicolumn{3}{l}{\textit{4. CES capital aggregator}} \\
& Output elasticity of public capital estimates & Bom \& Ligthart (2014) \\
\midrule
\multicolumn{3}{l}{\textit{5. Earnings-history pension}} \\
& Statutory replacement rates by earnings level & OECD Pensions at a Glance \\
& Minimum pension levels & National SS legislation \\
& Pension indexation rules & OECD / national sources \\
\midrule
\multicolumn{3}{l}{\textit{6. State-dependent income transitions}} \\
& Panel earnings data with health information & PSID / SHARE / EU-SILC \\
& Transition matrices estimated from panel data & Literature / own estimation \\
\bottomrule
\end{tabular}
\end{table}
